Thermal reactionome of a temperate ant species

### 2014-05-14 doc/ApTranscriptome_manuscript.Rmd
Updated manuscript with YAML header so can render with rmarkdown. Allows direction citation using pandoc rather than having to run separately.
Minor updates to text.

---
output:
  pdf_document:
    biblatex: yes
biblio-files: ApTranscriptome.bib
---
Thermal reactionome of a temperate ant species
============================================

**Authors**: John Stanton-Geddes, Andrew Nguyen, Mahesh Vangala, Jim Vincent, Rob Dunn, Aaron Ellison, Nate Sanders, Nick Gotelli, Sara Helms-Cahan

@@

The Trinity *de novo* transcriptome assembly included 126,172 genes with a total length of 100 million bp. Filtering to remove redundant or chimeric reads resulted in an assembly with 96,253 contigs and a total of 105,536 genes. For all genes, the total length was 63 million bp with an N~50~ length of 895 bp and a mean contig size of 593 bp. Of the 105,536 filtered genes, 55,432 had hits to the NCBI-nr database. Of these, 38,711 genes mapped to GO terms, 1,659 genes were identified to an enzyme and 18,935 genes mapped to a domain with >50% coverage. DeconSeq identified 5,675 genes as known contaminants from human, virus, bacterial or archael sources, leaving 99,861 clean transcripts. Of these, 47% have best hits to genes from the class Insecta.

@@

```{r expression_type_table, echo=FALSE, results='asis'}
pandoc.table(tt2, style="rmarkdown", caption = "Number of transcripts with maximum expression at high, low, intermediate, both high and low (bimodal) temperatures or are not thermally-responsivefor each species and their overlap.")
```

```{r expression_type_plot, echo=FALSE}
mosaicplot(tt2, shade = FALSE, xlab = "ApVT", ylab = "AcNC", main = "", cex = 0.5)
```

@@

![Probability density function of peak expression for all thermally-responsive in *A. carolinensis* (blue) and *A. picea* (red) species](../results/PDF_expression_all.png)

@@

@@

#![Expression (scaled to allow comparisons among genes) for genes in GO category "response to stress"](../results/GO_stress_High.png)

>>>>

---
title: "Thermal reactionome of a temperate ant species"
author: "John Stanton-Geddes, Andrew Nguyen, Mahesh Vangala, Jim Vincent, Rob Dunn,
  Aaron Ellison, Nate Sanders, Nick Gotelli, Sara Helms-Cahan"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  html_document: null
bibliography: ApTranscriptome.bib
csl: ecology.csl
---

## Abstract

When ants are exposed to different temperatures, lots of genes change expression.

@@

The Trinity *de novo* transcriptome assembly included 126,172 genes with a total length of 100 million bp. Filtering to remove redundant or chimeric reads resulted in an assembly with 105,536 genes. The total length was 63 million bp with an N~50~ length of 895 bp and a mean gene size of 593 bp. Of the 105,536 filtered genes, 55,432 had hits to the NCBI-nr database. Of these, 38,711 genes mapped to GO terms, 1,659 genes were identified to an enzyme and 18,935 genes mapped to a domain with >50% coverage. DeconSeq identified 5,675 genes as known contaminants from human, virus, bacterial or archael sources, leaving 99,861 clean transcripts. Of these, 47% have best hits to genes from the class Insecta [Supplementary Krona plot](http://johnstantongeddes.org/assets/files/ApTranscriptome_FA.html).

@@

```{r expression_type_table, echo=FALSE, results='asis'}
pandoc.table(tt2, style="rmarkdown", split.tables = Inf, caption = "Number of thermally-responsive genes in each expression category for each species.")
```

**Table 1 and Figure 1 are redundant - which is preferable??**

```{r expression_type_plot, echo=FALSE, fig.cap = "Mosaic plot that shows the proportion of thermally-responsive genes in each expression category for each species."}
mosaicplot(tt2, shade = FALSE, xlab = "ApVT", ylab = "AcNC", main = "", cex = 0.5)
```

@@

```{r exp_plot, echo=FALSE, warning = FALSE, fig.cap = "Probability density function of peak expression for all thermally-responsive in *A. carolinensis* (blue) and *A. carolinensis* (red) species"}

ggplot(Ap.df, aes(x=max.val, fill=colony)) +
  geom_density(alpha = 0.2, position = "identity") +
  scale_y_continuous(name = "Density") +
  scale_x_continuous(name = "Temperature of maximum expression")

```

@@

```{r T_on_plot, echo=FALSE, warning = FALSE, fig.cap = "Histogram with smooth density estimate of temperature of maximum change in expression for genes that have increased expression at *Low* and *High* temperatures. Bars are Probability density function of peak expression for all thermally-responsive in *A. carolinensis* (blue) and *A. carolinensis* (red) species"}
T_on_plot <- ggplot(Ap.T_on, aes(x = T_on, y = ..density.., colour = colony, group = colony, fill = colony)) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) + 
    scale_y_continuous(name="Density") +
    scale_x_continuous(name=expression("Temperature"))
print(T_on_plot)
```

@@
